---
title: "Marketing Campaign Analysis"
author: "Waranya Ratirotjanakul (36851256)"
date: today
format: 
    html: 
        theme: cosmo
        toc: true
        smooth-scroll: true
    pdf: 
        documentclass: article
        toc: true
---

```{r setup}
#| echo: false
#| warning: false
#| message: false

# load in all the used library
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
```

# Introduction 
Marketing expenditure plays a crucial role in influencing customer purchasing behaviour in online retail channels. It is important for businesses to understand whether increasing marketing investment leads to significant improvements in sales performance. Efficient allocation of marketing resources is essitial for maximise return on investment, supporting successful product launches, and sustaining long-term business growth. Therefore, this report investigates the relationship between marketing spending and customer purchase amount using campaign data.

# Research Question 
The report aims to investigate whether higher marketing expenditure per customer leads to increased purchase amounts in an online retail channel during the campaign period. The analysis focuses on identifying the relationship between marketing investment and customer purchasing behaviour across different customer segments, including enterprise, small and medium-sized businesses (SMBs), and consumers. The findings will help determine whether increasing marketing expenditure for personalised marketing campaigns is an effective strategy for improving revenue performance and support more efficient budget allocation.

# Dataset Introduction 
The marketing campaign dataset records customer interactions and purchasing behaviour during an online retail campaign. The dataset includes information on customer ID, customer registration date, marketing spending, number of visits, purchase amount, customer segment classification, and notes describing customer contact status (e.g., no response, contacted, and returned). 

Before analysis, the dataset was cleaned to ensure data accuracy and consistency. Placeholder values such as "N/A" and blank entries were identified as representations of missing data and were converted into proper missing values to enable consistent handling of incomplete information. In addition, the `signup_date` variable was converted from character format to a standard date format to ensure consistent representation of customer registration dates across the dataset.

```{r}
#| echo: false
#| warning: false
#| message: false

# load dataset and autometically convert "N/A" to real NA across column
mkt_campaign_raw <- read_csv("data/36851256_Waranya Ratirotjanakul.csv",
na = c("N/A", ""))

# convert character to date format
mkt_campaign_clean <- mkt_campaign_raw |>
    mutate(signup_date = parse_date_time(signup_date,
    orders = c("ymd", "dmy", "mdy")
    ),
    signup_date = as.Date(signup_date)
    )
```

# Dataset Description

## Data detail
The marketing campaign dataset contains `r nrow(mkt_campaign_clean)` observations and `r ncol(mkt_campaign_clean)` variables. The dataset includes numberic, categorical, character, and date-time variables. Numeric variables include `marketing_spend`, `visit`, and `purchase_amount`, while categoric variables includes customer `segment`. The variables `customer_id` and `notes` are stored as character data, and `signup_date` is recorded as a date variable.

## Data summary 
The summary statistics in @tbl-summarytable show differences in purchasing behaviour across segments in terms of marketing engagement and purchasing outcomes. Consumer customers represent the largest group (98 customers) and have the highest average purchase amount (169.84), indicating stronger overall spending compared with other segments. Enterprise customers show slightly higher average marketing spend (153.55) but a lower average purchase amount (140.95), suggesting that higher spending does not always translate directly into higher purchases. SMB customers have the lowest average purchase amount (122.50) despite moderate levels of marketing spend and visits.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summarytable
#| tbl-cap: Summary of Marketing Spend, Visits, and Purchase Amount by Customer Segment

# summary by segment
segment_summary <- mkt_campaign_clean |>
    group_by(segment) |>
    summarise(
        avg_marketing_spend = round(mean(marketing_spend, na.rm = TRUE), 2),
        avg_visits = round(mean(visits, na.rm = TRUE), 2),
        avg_purchase_amount = round(mean(purchase_amount, na.rm = TRUE), 2),
        total_customers = n()
    )

# display the table
segment_summary |>
    knitr::kable()
```

@fig-boxplot further illustrates the distribution of purchase amounts across customer segments. The boxplot shows that Consumer customers have both higher typical spending and greater variability in purchasing behaviour. Enterprise customers show a similar median purchase amount but a narrower spread, suggesting more consistent spending patterns within this segment. In contrast, SMB customers have a lower median purchase amount and a wider distribution, indicating greater variability and lower purchasing levels. These differences suggest that purchasing behaviour varies across customer segments and provide important context for evaluating marketing performance.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-boxplot
#| fig-cap: "Distribution of Purchase Amount by Customer Segment"

# create a summary plot
ggplot(mkt_campaign_clean,
       aes(x = segment,
           y = purchase_amount,
           fill = segment)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Purchase Amount by Customer Segment",
    x = "Customer Segment",
    y = "Purchase Amount"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Results 

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-spend
#| fig-cap: "Relationship Between Marketing Spend and Purchase Amount"

# create a scatter plot of marketing spend vs purchase amount
ggplot(mkt_campaign_clean,
       aes(x = marketing_spend,
           y = purchase_amount,
           colour = segment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm",
              color = "black",
              se = TRUE) +
  labs(
    x = "Marketing Spend",
    y = "Purchase Amount",
    colour = "Customer Segment"
  ) +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| results: false

# run linear regression model
model_perf <- lm(purchase_amount ~ marketing_spend,
                data = mkt_campaign_clean)

# evaluate model
summary(model_perf)
```

@fig-spend shows a strong positive relationship between marketing spend and purchase amount, indicating that customers who receive higher marketing investment tend to generate higher purchase values. The model explains approximately `r round(summary(model_perf)$r.squared * 100, 1)`% of the variation in purchase amount, suggesting that marketing spending is an important factor influencing customer purchasing behaviour.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-visit
#| fig-cap: "Relationship Between Number of Visits and Purchase Amount"

# create a scatter plot of visits vs purchase amount

ggplot(mkt_campaign_clean,
       aes(x = visits,
           y = purchase_amount,
           colour = segment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm",
              color = "black",
              se = TRUE) +
  labs(
    x = "Number of Visits",
    y = "Purchase Amount",
    colour = "Customer Segment"
  ) +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| results: false

# run linear regression model
model_visit <- lm(purchase_amount ~ visits,
                data = mkt_campaign_clean)

# evaluate model
summary(model_visit)
```

@fig-visit shows that the number of visits is also positively related to purchase amount, meaning customers who visit more frequently tend to spend more. However, the relationship appears weaker, with the model explaining only about `r round(summary(model_visit)$r.squared * 100, 1)`% of the variation in purchase amount. This suggests that while customer engagement contributes to purchasing behaviour, marketing spending plays a stronger role in driving purchase outcomes.

Although the results indicate a meaningful relationship between marketing spend and purchase amount, the scatter in the data shows that purchase amounts can still vary at similar levels of marketing spending. Therefore, collecting more campaign data in future periods would help improve the reliability of the analysis and support more confident evaluation of marketing effectiveness.