---
title: "Marketing Campaign Analysis"
author: "Waranya Ratirotjanakul (36851256)"
date: today
format: 
    html: 
        theme: cosmo
        toc: true
    pdf: 
        documentclass: article
        toc: true
---

```{r setup}
#| echo: false
#| warning: false
#| message: false

# load in all the used library
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
```

# Introduction 
Marketing expenditure plays a crucial role in influencing customer purchasing behaviour in online retail channels. It is important for businesses to understand whether increasing marketing investment leads to significant improvements in sales performance. Efficient allocation of marketing resources is essitial for maximise return on investment, supporting successful product launches, and sustaining long-term business growth. Therefore, this report investigates the relationship between marketing spending and customer purchase amount using campaign data.

# Research Question 
The report aims to investigate whether higher marketing expenditure per customer leads to increased purchase amounts in an online retail channel during the campaign period. The analysis focuses on identifying the relationship between marketing investment and customer purchasing behaviour across different customer segments, including enterprise, small and medium-sized businesses (SMBs), and consumers. The findings will help determine whether increasing marketing expenditure for personalised marketing campaigns is an effective strategy for improving revenue performance and support more efficient budget allocation.

# Dataset Introduction 
The marketing campaign dataset records customer interactions and purchasing behaviour during an online retail campaign. The dataset includes information on customer ID, customer registration date, marketing spending, number of visits, purchase amount, customer segment classification, and notes describing customer contact status (e.g., no response, contacted, and returned). 
Before analysis, the dataset was cleaned to ensure data accuracy and consistency. The `signup_date` variable was converted from character format to date-time format using `parse_date_time()` function to standardize date representation across the dataset. Missing values were inspected and verified to ensure data quality prior to analysis.

```{r}
# load dataset
mkt_campaign_raw <- read_csv("data/36851256_Waranya Ratirotjanakul.csv")

# inspect dataset
glimpse(mkt_campaign_raw)

# convert character to date-time format
mkt_campaign_clean <- mkt_campaign_raw |>
    mutate(signup_date = parse_date_time(signup_date,
    orders = c("ymd", "dmy", "mdy")
    ))
glimpse(mkt_campaign_clean)

# check missing value
mkt_campaign_clean |>
  filter(is.na(signup_date))
mkt_campaign_raw |>
  filter(is.na(signup_date))
```

# Dataset Description

## Data detail
The marketing campaign dataset contains `r nrow(mkt_campaign_clean)` observations and `r ncol(mkt_campaign_clean)` variables. The dataset includes numberic, categorical, character, and date-time variables. Numeric variables include `marketing_spend`, `visit`, and `purchase_amount`, while categoric variables includes customer `segment`. The variables `customer_id` and `notes` are stored as character data, and `signup_date` is recorded as a date-time variable.

## Data summary 
The scatter plot shows a positive relationship between marketing spend and purchase amount, indicating that customers who receive higher marketing investment tend to generate higher purchase values. However, some variation is observed across customers, suggesting that additional factors may also influence purchasing behaviour.

```{r}
#| fig-cap: "Relationship Between Marketing Spend and Purchase Amount"

# create a summary plot marketing spend vs purchase amount
ggplot(mkt_campaign_clean,
       aes(x = marketing_spend,
           y = purchase_amount)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm") +
  labs(
    x = "Marketing Spend",
    y = "Purchase Amount"
  ) +
  theme_minimal()

# run linear regression model
model_perf <- lm(purchase_amount ~ marketing_spend,
                data = mkt_campaign_clean)

# evaluate model
summary(model_perf)
```

The table shows differences in average purchasing behaviour across customer segments. Consumer customers have the highest average purchase amount, while SMB customers show relatively lower spending levels. These differences provide context for understanding how marketing performance may vary across customer groups.

```{r}
# summary by segment
segment_summary <- mkt_campaign_clean |>
    group_by(segment) |>
    summarise(
        avg_marketing_spend = round(mean(marketing_spend, na.rm = TRUE), 2),
        avg_visits = round(mean(visits, na.rm = TRUE), 2),
        avg_purchase_amount = round(mean(purchase_amount, na.rm = TRUE), 2),
        total_customers = n()
    )

# display the table
segment_summary |>
    knitr::kable(
        caption = "Table1: Summary of Marketing Spend, Visits, and Purchase Amount by Customer Segment"
    )
```

```{r}
# create a summary plot
ggplot(mkt_campaign_clean,
       aes(x = segment,
           y = purchase_amount)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Purchase Amount by Customer Segment",
    x = "Customer Segment",
    y = "Purchase Amount"
  ) +
  theme_minimal()
```

# Results 

A simple linear regression model was used to examine the relationship between marketing spend and purchase amount. The results indicate a positive and statistically significant relationship between the two variables (β = 0.896, p < 0.001). This suggests that higher marketing spending is associated with increased customer purchase amounts. The model explains approximately 74.1% of the variation in purchase amount (R² = 0.741), indicating a strong relationship between marketing investment and purchasing behaviour. A total of 18 observations were excluded from the analysis due to missing values in the dataset.

