---
title: "Marketing Campaign Analysis"
author: "Waranya Ratirotjanakul (36851256)"
date: today
format: 
    html: 
        theme: cosmo
        toc: true
        smooth-scroll: true
    pdf: 
        documentclass: article
        toc: true
---

```{r setup}
#| echo: false
#| warning: false
#| message: false

# load all required libraries
library(knitr)
library(kableExtra)
library(tidyverse)
library(lubridate)
library(ggplot2)
```

# Introduction 
Marketing expenditure plays a crucial role in influencing customer purchasing behaviour in online retail channels. Understanding whether increased marketing investment leads to improvements in sales performance is essential for effective business decision-making. Efficient allocation of marketing resources helps maximise return on investment, support successful product launches, and sustain long-term business growth. Therefore, this report investigates the relationship between marketing spending and customer purchase amount using campaign data.

# Research Question 
This report investigates whether higher marketing expenditure per customer leads to increased purchase amounts in an online retail channel during the campaign period. The analysis examines the relationship between marketing investment and customer purchasing behaviour across different customer segments, including enterprise, small and medium-sized businesses (SMBs), and consumers. The findings will help determine whether increasing marketing expenditure for personalised marketing campaigns is an effective strategy for improving revenue performance and supporting more efficient budget allocation.

# Dataset Introduction 
The marketing campaign dataset records customer interactions and purchasing behaviour during an online retail campaign. It includes information on customer ID, customer registration date, marketing spending, number of visits, purchase amount, customer segment classification, and notes describing customer contact status (e.g., no response, contacted, and returned). This data provides insight into customer engagement and purchasing outcomes associated with marketing activities.

Before analysis, the dataset was cleaned to ensure data accuracy and consistency. Placeholder values such as "N/A" and blank entries were identified as representations of missing data and were converted into proper missing values to enable consistent handling of incomplete information. In addition, the `signup_date` variable was converted from character format into a standard date format to ensure consistent representation of customer registration dates across the dataset.

```{r}
#| echo: false
#| warning: false
#| message: false

# load dataset and automatically convert "N/A" and blank to real NA across column
mkt_campaign_raw <- read_csv(
    "data/36851256_Waranya Ratirotjanakul.csv",
    na = c("N/A", "")
)

# convert character to date format
mkt_campaign_clean <- mkt_campaign_raw |>
    mutate(
        signup_date = parse_date_time(
            signup_date,
            orders = c("ymd", "dmy", "mdy")
        ),
        signup_date = as.Date(signup_date)
    )
```

## Dataset Description
The marketing campaign dataset contains `r nrow(mkt_campaign_clean)` observations and `r ncol(mkt_campaign_clean)` variables. It includes numeric, categorical, character, and date variables representing customer behaviour and marketing activity. Numeric variables include `marketing_spend`, `visits`, and `purchase_amount`, while the categorical variable includes customer `segment`. The variables `customer_id` and `notes` are stored as character data, and `signup_date` is recorded as a date variable.

## Dataset summary 
The summary statistics in @tbl-summarytable show differences in purchasing behaviour across segments in terms of marketing engagement and purchasing outcomes. Consumer customers represent the largest group (98 customers) and have the highest average purchase amount (169.84), indicating stronger overall spending compared with other segments. Enterprise customers show slightly higher average marketing spend (153.55) but a lower average purchase amount (140.95), suggesting that higher spending does not always translate into higher purchases. SMB customers have the lowest average purchase amount (122.50) despite moderate levels of marketing spend and visits, indicating relatively weaker purchasing behaviour within this segment.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summarytable
#| tbl-cap: Summary of Marketing Spend, Visits, and Purchase Amount by Customer Segment

# summary by segment
segment_summary <- mkt_campaign_clean |>
    group_by(segment) |>
    summarise(
        avg_marketing_spend = round(mean(marketing_spend, na.rm = TRUE), 2),
        avg_visits = round(mean(visits, na.rm = TRUE), 2),
        avg_purchase_amount = round(mean(purchase_amount, na.rm = TRUE), 2),
        total_customers = n()
    )

# display the table
segment_summary |>
    knitr::kable()
```

The boxplot in @fig-boxplot further illustrates the distribution of purchase amounts across customer segments. Consumer customers show both higher typical spending and greater variability in purchasing behaviour. Enterprise customers show a similar median purchase amount but a narrower spread, suggesting more consistent spending patterns within this segment. In contrast, SMB customers have a lower median purchase amount and a wider distribution, indicating greater variability and lower purchasing levels. These differences suggest that purchasing behaviour varies across customer segments and provide important context for evaluating marketing performance.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-boxplot
#| fig-cap: "Distribution of Purchase Amount by Customer Segment"

# create a summary plot
ggplot(mkt_campaign_clean,
       aes(x = segment,
           y = purchase_amount,
           fill = segment
        )
) +
  geom_boxplot() +
  labs(
    title = "Distribution of Purchase Amount by Customer Segment",
    x = "Customer Segment",
    y = "Purchase Amount"
    ) +
  theme_minimal() +
  theme(
    legend.position = "none"
    )
```

# Results 
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-spend
#| fig-cap: "Relationship Between Marketing Spend and Purchase Amount"

# create a scatter plot of marketing spend vs purchase amount
ggplot(mkt_campaign_clean,
       aes(x = marketing_spend,
           y = purchase_amount,
           colour = segment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm",
              color = "black",
              se = TRUE) +
  labs(
    x = "Marketing Spend",
    y = "Purchase Amount",
    colour = "Customer Segment"
  ) +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| results: false

# run linear regression model
model_perf <- lm(
    purchase_amount ~ marketing_spend,
    data = mkt_campaign_clean
)

# evaluate model
summary(model_perf)
```

The scatter plot in @fig-spend shows a strong positive relationship between marketing spend and purchase amount, indicating that customers who receive higher marketing investment tend to generate higher purchase values. The regression model explains approximately `r round(summary(model_perf)$r.squared * 100, 1)`% of the variation in purchase amount, demonstrating that marketing spending is a meaningful predictor of customer purchasing behaviour. This suggests that increasing marketing expenditure is associated with higher purchase amounts during the campaign period.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-visit
#| fig-cap: "Relationship Between Number of Visits and Purchase Amount"

# create a scatter plot of visits vs purchase amount
ggplot(mkt_campaign_clean,
       aes(x = visits,
           y = purchase_amount,
           colour = segment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm",
              color = "black",
              se = TRUE) +
  labs(
    x = "Number of Visits",
    y = "Purchase Amount",
    colour = "Customer Segment"
  ) +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| results: false

# run linear regression model
model_visit <- lm(
    purchase_amount ~ visits,
    data = mkt_campaign_clean
)

# evaluate model
summary(model_visit)
```

The relationship between customer engagement and purchasing behaviour is further illustrated in @fig-visit. Customers who visit more frequently also tend to spend more, although the relationship appears weaker, with the model explaining approximately `r round(summary(model_visit)$r.squared * 100, 1)`% of the variation in purchase amount. This indicates that while customer engagement contributes to purchasing behaviour, marketing spending plays a stronger role in influencing purchase outcomes.

Overall, the analysis provides clear evidence that higher marketing expenditure per customer is associated with increased purchase amounts. However, the variability observed in the data suggests that marketing spending alone does not fully explain purchasing behaviour, and the current dataset size (150 customers) may be insufficient to capture the full range of customer responses. Collecting additional campaign data, including a larger number of customers and observations from future campaign periods, would help improve the reliability and applicability of the analysis. Furthermore, incorporating more detailed customer information, such as demographics, purchase history, and marketing channel preferences, would support more accurate modelling and more data-driven decision-making in marketing planning.